\SSbreak\\
\emph{Source: AIME, 1985 P14}\\
\emph{Proposer: \Pchan}\\
\emph{Problem ID: 79}\\
\emph{Date: 2020-12-25}\\
\SSbreak

\SSpsetQ{
    Santa has a number of elves. He wants to select the very best for Christmas. To do so, he makes all his elves compete in a \sout{present wrapping competition} War Thunder tournament. Each elf competes against every other elf, and the winner receives one point, while the loser receives no points. If a particular match turns out to be a draw, each elf is awarded 0.5 points. \medskip

    After the tournament has finished, Santa notices that exactly half of the points earned by each elf was done so when matched against the ten lowest scoring elves.\medskip
    
    Given that each of the ten lowest scoring elves also earned half of their points against the other nine, how many elves does Santa have to choose from?
}\bigskip

\begin{solution}\hfil\medskip
    
    Let there be $n+10$ elves. Then the bottom $10$ elves score ${10 \choose 2}$ elves from each other while the rest score ${n \choose 2}$ points from those that are not in the bottom $10$ people, and thus ${n \choose 2}$ points from the bottom $10$ elves. Thus we get $\frac{1}{2} {n+10 \choose 2} = {n \choose 2} +{10 \choose 2}$ and thus solving we get $n = 6,15$, but if $n= 6$, we get that it is expected for a player from the bottom $10$ players will win when matched  against a player from the top $6$, which is impossible. Thus, $n=15$ and there are $25$ elves Santa can choose from.
\end{solution}\bigskip
