\SSbreak\\
\emph{Source: IZhO 2021/5}\\
\emph{Proposer: \Phyper}\\ %\Pchan \Pbrain \Pss
\emph{Problem ID: 188}\\
\emph{Date: 2021-05-02}\\
\emph{Difficulty: Challenging}\\
\SSbreak

\SSpsetQ{
On a party with $102$ guests, hosts Tony and Wang play a game (the hosts are not regarded as guests). There are $102$ chairs arranged in a circle; initially, all guests hang around those chairs. The hosts take turns alternately. By a turn, a host orders any standing guest to sit on an unoccupied chair $c$. If some chair adjacent to $c$ is already occupied, the same host orders one guest on such chair to stand up (if both chairs adjacent to $c$ are occupied, the host chooses exactly one of them). All orders are carried out immediately. Tony makes the first move; her goal is to fulfill, after some move of hers, that at least $k$ chairs are occupied. Determine the largest $k$ for which Tony can reach the goal, regardless of Wang's play.

%Put Problem Here
}\bigskip

\begin{solution}\hfil\medskip
 
We claim the answer is \fbox{35}. Note the number of chairs does not decrease, given that every empty chair must be next to a full chair, we have at least 34 chairs. In addition, when there are 33 chairs, every third chair is occupied.
 
 If Wang forces Tony into such a configuration, say Tony placed a guest on chair $x$ and Wang on $y$. However, if $y$ is adjacent to $x$, Tony could have easily have chosen $y$ instead and wins. Otherwise, she could have placed it on $x\pm 1$, as there were no guests there, and now not every third chair is occupied. Thus, Wang never reaches a position with 34 chairs such that no more guests can be placed for Tony, and Tony can happily place her last guest.
 
 To show that Wang can win otherwise, we show inductively that Wang can hold a strategic position:\\
 \textbf{Claim.} Wang can make sure that in the sets $S_2,S_3,\ldots,S_{34}$ where $S_k=\{3k+1,3k+2,3k+3\}$, none of $3k+1$ and $3k+3$ is occupied. Furthermore, there is at most two occupied chairs in $S_1$.\\
 \textbf{Proof.} The proof is by induction.\\
 \textit{Base Case.} No moves are made, in which case the claim holds.\\
 \textit{Induction Hypothesis.} Assume that Wang holds such a position. We claim after 2 moves, Wang can hold this position again.
 \textit{Induction Step.} If Alice's move holds the position, Wang can simply put a person somewhere in $S_1$ (as chairs 4 and 102 are unoccupied, we can once again only have at most 2 people in $S_1$). If it doesn't, she could not have operated on $S_1$, and thus placed a guest in some of $S_2,\ldots,S_{99}$. However, if she operates on $S_k$, Wang places a person in the middle of $S_k$ and removes the mischief caused by Alice. Thus Wang can hold his position.
%Put sol here
\end{solution}\bigskip
