\SSbreak\\
\emph{Source: 2019 China Mathematical Olympiad P1, Modified}\\
\emph{Proposer: \Paiya}\\ %\Pchan \Pbrain \Pss
\emph{Problem ID: 174}\\
\emph{Date: 202-04-11}\\
\emph{Difficulty: Challenging}\\
\SSbreak

\SSpsetQ{
	Let $\left(a_1, a_2, a_3, a_4, a_5, a_6\right)$ be reals no less than $-2020$ that sum to $2021$. 
    Let $M$ and $m$ be the maximum and minimum, respectively, of the function 
    $$f\left(a_1, a_2, a_3, a_4, a_5, a_6\right) = \prod_{k = 1}^6 \left(a_k + a_{k + 1}\right),$$ where indices are taken modulo $6$,
    and let $N$ and $n$ be the number of ordered pairs which maximize and minimize $f$, respectively.
    Find $\left\lfloor \ln(-Mm) \right\rfloor + N + n$. \medskip

    \begin{center}
        \textit{(A scientific calculator may be used.)}
    \end{center}
	%Put Problem Here
}\bigskip

\begin{solution}\hfil\medskip
	
	We switch variables to $(a, b, c, d, e, f)$ for sake of legibility. WLOG $a$ is maximal; therefore $a + b, f + a \geq 0$. 
    We solve for the minimum first, as that is less painful. Negative values are possible, which correspond to either one or three negative factors.
    If there is only one negative factor, then there are at least two positive variables less than $2021$; flipping their signs and increasing
    the value of the greatest two variables accordinaly produces a product greater in absolute value, so we want three negative factors.
    Let $a + b, b + c, c + a < 0$ (the factor adjacency really does matter as we'll see in a bit). Applying AM-GM twice, once to the positive factors and once to the negative, we have
    $$f \leq \left(\dfrac{-a - 2b - 2c - d}{3}\right)^3\left(\dfrac{d + 2e + 2f + a}{3}\right)^3$$
    and since $d + 2e + 2f + a = 2 \cdot 2021 - (a + 2b + 2c + d)$ we see that the maximum occurrs when $a + 2b + 2c + d$ is minimized;
    this happens at $a = b = c = d = -2020$ and $e = f = \frac{10101}{2}$. Note that the variables with coefficient $2$ must be different variables; in other words, the three factors we chose to be negative matters.
    In esssence, the two positive terms have to be next to each other, so $n = 6$.\medskip

    Now for the maximum. By the same argument, having all positive factors does not yield a maximal configuration; therefore there are either two or four negative factors.
    Applying AM-GM in the same fashion (once to the positive factors and once to the negative factors) we get that the equality case is one of four
    and that the maximal configuration happens when all the variables in the negative terms are equal to $-2020$. Thus, four negative factors is optimal;
    this happens at $a = b = c = d = e = -2020$ and $f = 12121$ and $N = 6$.\medskip

    Our answer is $$\left\lfloor \ln\left(2^5 \cdot 2020^6 \cdot 6061^2 \cdot 10101^3\right) \right\rfloor + 6 + 6 = \boxed{113}.$$
	%Put sol here
\end{solution}\bigskip
