\SSbreak\\ 
\emph{Source: \Cop}\\
\emph{Proposer: \Pbfan}\\
\emph{Problem ID: 31}\\
\emph{Date: 2020-11-02}\\
\SSbreak

\SSpsetQ{
    Tan and Wen are writing questions for CCCC at a rate of 1 per minute. Immediately after Tan
     writes a question, Wen eats his paper with probability \( \frac 17 \), so that Tan must restart.\\

    After an extremely long time (assume infinite), Tony Wang walks in.
     What is the expected number of questions Tony Wang sees written on Tan's paper?
}\bigskip 

\begin{solution}\hfil\medskip

    Let $E_n$ be the expected number of questions Tony Wang sees written on Tan's paper after $n$ minutes. \\

    Then the following recurrence holds: 
    \[E_{n+1} = \frac{6}{7} (E_n+1) + \frac{1}{7} \cdot 0\]
    because with $\frac{6}{7}$ probability, Tan writes another question without Wen eating it, 
    and with $\frac{1}{7}$ probability, Wen eats all of Tan's questions. \\

    Claim 1: $E_n < 6 ~\forall n$. 

    \begin{subproof}
        The proof is by induction. Note $E_0 = 0$. If $E_k < 6$, then 
        $E_{k+1} = \frac{6}{7} (E_k + 1) < \frac{6}{7} (6 + 1) = 6$. 
        This completes the proof. 
    \end{subproof}

    Claim 2: $E_{n+1} > E_n ~\forall n$. 

    \begin{subproof}
        Note that since $E_k < 6 ~\forall k$, we have 
        $E_{k+1} = \frac{6}{7} (E_k+1) > \frac{6}{7} (E_k + \frac16 E_k) = E_k$. 
    \end{subproof}

    Now by the Monotone Bounded Convergence Theorem, the sequence $\left( E_n \right )$ converges to a limit. 
    Suppose $\lim E_n = L$. Then note $\lim E_{n+1} = L$ since changing the first terms of a sequence does not change the overall convergence. 
    So since $E_{n+1} = \frac{6}{7} (E_n + 1)$, $\lim E_{n+1} = \lim \frac{6}{7} (E_n+1)$. 
    So $L = \frac{6}{7}(L + 1)$. \\

    Solving this, we obtain $L = \boxed{6}$. 
\end{solution}\bigskip