To start off, we’d like to apologise for the noticeable drop in quality for the past two seasons (Trigonometric Troubles \& Piboi’s 14 Combi Problems). This has mostly been a result of us being busy with other things at the moment (university interviews, finals, etc.) Consequently, we have not had the time to properly plan seasons in advance, and in the case of the last season, test-solve. In conjunction with us not being able to put in the time, we got carried away in terms of focusing on things like flavour-text and themed seasons. We can see how this may have alienated people who were confused by a particular problem statement, didn’t enjoy the topic of the season, and so forth. We will try to eliminate these issues by planning seasons ahead, and in more detail. However, to do this, there may be times when breaks happen in between seasons. Furthermore, from now on we plan on keeping the seasons varied topic-wise, and any seasons of a thematic nature will also have varied problems.\\

On what happened this season: We would like to say we completely dropped the ball with it, and in no way is Piboi responsible for that. As aforementioned, the quality of this season dropped so much mostly due to us not having the time to sort things out beforehand, and thus, we did not rigorously review the questions and solutions. This was particularly evident in the first question of the season, and again today. After careful deliberation, we have decided to void this season and start again with a new one starting on Monday, we will be less negligent in future.\\

Finally, We appreciate all of the support, and are deeply sorry that there have been many issues over the past week or two, and hope that you continue to support the QOTD's. We will do our best to prevent the issues which have cropped up in the past two weeks, and in doing so improve the quality of the QOTD’s so it’s enjoyable for everyone.\\ 

We’ll be opening up a little consultation over the weekend where you can give us some suggestions on any changes you’d perhaps want to see for future seasons, be it technical changes or changes to the format, we want to hear it. We’ve added a channel on the OpenPOTD server where you can add them - alternatively you can DM any of us.\\

Thank you for reading, sincerely \\

@brainysmurfs\#2860  @.sjbs\#9839 @Angry Any\#4319 \\

\url{https://discord.gg/GsPSSHdhPB}